Parameter Estimation for the Compound Poisson Model

### 2017-01-09 reports/compoundPoisson/compoundPoisson.tex
compoundPoisson report, spell check.

For low $\nu\tau$, there is a chance of simulating zeros. The Normal approximation failed to capture the probability mass at zero because the Normal distribution was constrained to be symmetric. The saddlepoint approximation did capture probabilty mass at zero as shown by an increase in probability density towards zero. However for a distinct mixture looking compound Poisson, Figure \ref{fig:histogram} (d), the saddlepoint approxmation was not flexible enough to capture each individual peaks.

For the Normal looking compound Poisson, both approxmations performed well.
\begin{figure*}
\centerline{
	\centering
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/normal_1.eps}
        \caption{Normal approximation: $\nu\tau=1$, $\alpha=1$, $\lambda=1$}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/saddle_1.eps}
        \caption{Saddlepoint approximation: $\nu\tau=1$, $\alpha=1$, $\lambda=1$}
    \end{subfigure}
}
\centerline{
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/normal_2.eps}
        \caption{Normal approximation: $\nu\tau=3$, $\alpha=400$, $\lambda=100$}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/saddle_2.eps}
        \caption{Saddlepoint approximation: $\nu\tau=3$, $\alpha=400$, $\lambda=100$}
    \end{subfigure}
}
\centerline{
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/normal_3.eps}
        \caption{Normal approximation: $\nu\tau=500$, $\alpha=80$, $\lambda=2$}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/saddle_3.eps}
        \caption{Saddlepoint approximation: $\nu\tau=500$, $\alpha=80$, $\lambda=2$}
    \end{subfigure}
}
\caption{A sample of $10^4$ compound Poisson random variables was simulated. The histogram of the samples was plotted along with the density approxmation.}
\label{fig:histogram}
\end{figure*}

\section{Parameter estimation}
The saddlepoint approxmation did capture the probabilty mass at zero, however the formula for the density approxmation does not support zero. As a result, the saddlepoint approxmation may not be suited to approximate the likelihood when there is a good chance of obtaining zeros in the data. A way around it is to adjust the data by adding a very small constant to all zeros in the data.

>>>>

For low $\nu\tau$, there is a chance of simulating zeros. The Normal approximation failed to capture the probability mass at zero because the Normal distribution was constrained to be symmetric. The saddlepoint approximation did capture probability mass at zero as shown by an increase in probability density towards zero. However for a distinct mixture looking compound Poisson, Figure \ref{fig:histogram} (d), the saddlepoint approximation was not flexible enough to capture each individual peaks.

For the Normal looking compound Poisson, both approximation performed well.
\begin{figure*}
\centerline{
	\centering
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/normal_1.eps}
        \caption{Normal approximation: $\nu\tau=1$, $\alpha=1$, $\lambda=1$}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/saddle_1.eps}
        \caption{Saddlepoint approximation: $\nu\tau=1$, $\alpha=1$, $\lambda=1$}
    \end{subfigure}
}
\centerline{
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/normal_2.eps}
        \caption{Normal approximation: $\nu\tau=3$, $\alpha=400$, $\lambda=100$}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/saddle_2.eps}
        \caption{Saddlepoint approximation: $\nu\tau=3$, $\alpha=400$, $\lambda=100$}
    \end{subfigure}
}
\centerline{
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/normal_3.eps}
        \caption{Normal approximation: $\nu\tau=500$, $\alpha=80$, $\lambda=2$}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/saddle_3.eps}
        \caption{Saddlepoint approximation: $\nu\tau=500$, $\alpha=80$, $\lambda=2$}
    \end{subfigure}
}
\caption{A sample of $10^4$ compound Poisson random variables was simulated. The histogram of the samples was plotted along with the density approxmation.}
\label{fig:histogram}
\end{figure*}

\section{Parameter estimation}
The saddlepoint approximation did capture the probability mass at zero, however the formula for the density approximation does not support zero. As a result, the saddlepoint approxmation may not be suited to approximate the likelihood when there is a good chance of obtaining zeros in the data. A way around it is to adjust the data by adding a very small constant to all zeros in the data.
